---
title: "Initial Korean covid 19"
author: "Colton Beardsley"
date: "10/10/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?


# Method

Based on the data we found in kRdata, we will try to find out if there's a difference between men and women. Is there any difference in how COVID kills men and women? Is COVID more harmful to men? Or is it more harmful to women? 

This appears to be a category-category problem. Therefore, we will be deploying cat-cat helper files to run a test. 


## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Looking at the first chart, it shows that more women got COVID then men.

When looking at the second graph, the results show that of the whole group of people that got COVID, more men were being killed than women.




  
## Numerical results

```{r}
table2 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```

The result in the data here confirms the data that was present in the ggplot results, which was that women got COVID at a higher rate then men. However, it also shows that of the people who got COVID, more men died than women.


## Inferential results 


```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)

```


The results of the chrisquare and fisher test show that men are twice as likely to die from COVID than women.

## Conclusions

We first looked at the kRdata and confirmed that this was a category~category problem. We then used our Cat~Cat helper files to run a ggplot test. From the results of the ggplot, we determined that of the people who get COVID, women appear to get COVID at a higher rate. It also seems to appear that of the people who got COVID, men died at a higher rate then women. However, we needed to run more test to confirm.

